---
title: "About"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{About}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Citing scopusflow

If you use scopusflow in published work, please cite it.

> Bernabeu, P. (2026). scopusflow: A reproducible workflow layer for 'Scopus'
> bibliographic searches. R package version `r packageVersion("scopusflow")`.
> https://doi.org/10.5281/zenodo.21252669

```{r bibtex, echo = FALSE, results = "asis"}
# Build the BibTeX entry from the installed version so it never drifts, then
# render it with a copy button and a download link. The button uses the
# browser clipboard API; the link is a self-contained data URI, so neither
# depends on a static file being shipped alongside the site.
ver <- as.character(utils::packageVersion("scopusflow"))
bib <- paste(
  "@Manual{scopusflow,",
  "  title  = {{scopusflow}: A reproducible workflow layer for '{Scopus}' bibliographic searches},",
  "  author = {Pablo Bernabeu},",
  "  year   = {2026},",
  sprintf("  note   = {R package version %s},", ver),
  "  doi    = {10.5281/zenodo.21252669},",
  "  url    = {https://doi.org/10.5281/zenodo.21252669},",
  "}",
  sep = "\n"
)
esc <- function(x) {
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  gsub(">", "&gt;", x, fixed = TRUE)
}
uri <- paste0(
  "data:application/x-bibtex;charset=utf-8,",
  utils::URLencode(bib, reserved = TRUE)
)
cat(sprintf(
'<div class="citation-bibtex">
<pre id="scopusflow-bibtex"><code>%s</code></pre>
<p class="citation-bibtex-actions">
<button type="button" class="btn btn-primary btn-sm" onclick="scopusflowCopyBibtex(this)">Copy BibTeX</button>
<a class="btn btn-outline-primary btn-sm" download="scopusflow.bib" href="%s">Download .bib</a>
</p>
</div>
<script>
function scopusflowCopyBibtex(btn) {
  var code = document.getElementById("scopusflow-bibtex");
  navigator.clipboard.writeText(code.innerText).then(function () {
    var label = btn.textContent;
    btn.textContent = "Copied";
    setTimeout(function () { btn.textContent = label; }, 1500);
  });
}
</script>
', esc(bib), uri))
```

R users can also retrieve this citation directly with `citation("scopusflow")`.

## The developer

[Pablo Bernabeu](https://pablobernabeu.github.io/) is a researcher in the
Department of Education at the University of Oxford, with hands-on experience
of behavioural experiments, EEG, corpus analysis, computational modelling and
statistics. He develops open, reproducible research software in R and Python,
and is a Fellow of the Software Sustainability Institute. scopusflow and its
[Python twin](https://pablobernabeu.github.io/scopusflow-py/) are part of that
work, keeping a search reproducible and its results legible across both
languages.

## Licence

scopusflow is released under the [MIT licence](../LICENSE.html). 'Scopus' is a
trademark of Elsevier. scopusflow is an independent client and is not
affiliated with or endorsed by Elsevier.

## Versioning and archival

Releases are tagged on GitHub and archived on Zenodo. The concept DOI
[10.5281/zenodo.21252669](https://doi.org/10.5281/zenodo.21252669) always
resolves to the latest archived version, so a citation stays current without
naming a version. The [changelog](../news/index.html) records what changed in
each release.

## Contributing and support

Bugs and feature requests are best raised on the
[GitHub issues page](https://github.com/pablobernabeu/scopusflow/issues), and
the [contributing guide](../CONTRIBUTING.html) describes the development setup
and the conventions the package follows. Because the package works with an
Elsevier API key, please never paste a key or any other secret into an issue.
