---
title: "Analysing and visualising a literature"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysing and visualising a literature}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
# Render data frames and tibbles as tables rather than console text.
local({
  kp <- function(x, ...) {
    if (any(vapply(x, is.list, logical(1)))) return(knitr::normal_print(x))
    knitr::knit_print(knitr::kable(x))
  }
  for (cls in c("data.frame", "tbl_df", "tbl")) {
    registerS3method("knit_print", cls, kp, envir = asNamespace("knitr"))
  }
})
has_ggplot2 <- requireNamespace("ggplot2", quietly = TRUE)
```

```{r setup}
library(scopusflow)
```

Once a set of records is in hand, the package offers a small analysis layer that
turns it into the figures a bibliometric study usually needs. The steps that
contact the API are shown but not run; the rest run offline, on synthetic
records built in the shape the API returns.

## What is in a record set

`scopus_top()` tallies the most frequent sources or authors. Author strings that
hold several names are split, so each contributor is counted once per record.
The bundled `example_records` keeps the package's examples quick, but its ten
single-appearance records make for a flat tally, so here we synthesise a
slightly larger corpus and normalise it with `scopus_records()`, exactly as a
retrieval would.

```{r}
lead <- rep(c("Chen L.", "Smith J.", "Garcia M.", "Kumar S.", "Tanaka H.", "Okafor N."),
            times = c(8, 6, 4, 4, 2, 2))
raw <- list(entry = lapply(seq_along(lead), function(i) list(
  `dc:identifier` = paste0("SCOPUS_ID:", 85100000000 + i),
  `dc:title` = sprintf("Synthetic study %02d", i),
  `dc:creator` = if (i %% 3 == 0) paste(lead[i], "Novak P.", sep = "; ") else lead[i],
  `prism:publicationName` = rep(c("Nature", "Science", "Physical Review Letters",
                                  "Advanced Materials", "The Lancet Oncology"),
                                times = c(9, 7, 5, 3, 2))[i],
  `prism:coverDate` = sprintf("%d-06-01", rep(2016:2023, times = c(1, 1, 2, 2, 3, 4, 6, 7))[i]),
  `citedby-count` = as.character((27 - i) * 3)
)))
records <- scopus_records(raw, query = "TITLE-ABS-KEY(synthetic corpus)")

scopus_top(records, by = "source")
scopus_top(records, by = "author", n = 5)
```

```{r eval = has_ggplot2, fig.alt = "A horizontal bar chart of the most frequent sources", fig.width = 7, fig.height = 3.5}
plot_scopus_top(scopus_top(records, by = "source"))
```

The same plot works on the author tally.

```{r eval = has_ggplot2, fig.alt = "A horizontal bar chart of the most frequent authors", fig.width = 7, fig.height = 3.5}
plot_scopus_top(scopus_top(records, by = "author", n = 5))
```

A record set also has an honest default view: `autoplot()` draws its records per
year. The same `autoplot()` generic dispatches on `scopus_trend` and `scopus_top`
objects too, delegating to the plots above.

```{r eval = has_ggplot2, fig.alt = "A bar chart of records per year, rising over time", fig.width = 7, fig.height = 3.5}
ggplot2::autoplot(records)
```

## How a literature grows

`scopus_trend()` counts how many records match a query in each year, which is the
size of a literature over time. It issues one count request per year, so it needs
the API.

```{r eval = FALSE}
tr <- scopus_trend("graphene", years = 2004:2022, field = "TITLE-ABS-KEY")
plot_scopus_trend(tr)
```

The result has a fixed shape, which we reproduce here to show the plot.

```{r eval = has_ggplot2, fig.alt = "A line and area chart of records per year rising over time", fig.width = 7.5, fig.height = 4}
years <- 2004:2022
tr <- tibble::tibble(query = "TITLE-ABS-KEY(graphene)", year = years,
                     n = round(exp(seq(log(50), log(28000), length.out = length(years)))))
class(tr) <- c("scopus_trend", class(tr))
plot_scopus_trend(tr)
```

## Where a niche sits

A study that crosses two or more fields is best introduced by sizing those
fields and their overlap: each parent literature may hold thousands of records
while their intersection holds a handful, which is the niche the study
occupies. `scopus_intersections()` counts a named set of concepts and any
requested intersections of them, at one count request per row, so the whole
landscape is as cheap as a few `scopus_count()` calls. Concept values may be
bare terms, wrapped in `field` for you, or complete field-tagged expressions,
used exactly as given.

```{r eval = FALSE}
sets <- scopus_intersections(
  concepts = c(
    "semantic priming"  = "semantic priming",
    "mental simulation" = "mental simulation"
  ),
  intersections = list(c("semantic priming", "mental simulation")),
  field = "TITLE-ABS-KEY"
)
plot_scopus_intersections(sets, highlight = sets$label[sets$type == "intersection"])
```

As above, the result has a fixed shape, which we reproduce here to show the
plot. The lollipop chart uses a log-scale axis, so the small intersection stays
legible beside its large parent fields, and the highlighted row draws the eye
to the niche itself.

```{r eval = has_ggplot2, fig.alt = "A log-scale lollipop chart showing two large concepts and their small intersection, with the intersection highlighted", fig.width = 7.5, fig.height = 2.6}
sets <- tibble::tibble(
  label = c("semantic priming", "mental simulation",
            "semantic priming × mental simulation"),
  query = c("TITLE-ABS-KEY(semantic priming)",
            "TITLE-ABS-KEY(mental simulation)",
            "(TITLE-ABS-KEY(semantic priming)) AND (TITLE-ABS-KEY(mental simulation))"),
  n = c(6600, 2100, 15),
  type = c("concept", "concept", "intersection"),
  size = c(1L, 1L, 2L),
  members = c("semantic priming", "mental simulation",
              "semantic priming; mental simulation")
)
class(sets) <- c("scopus_intersections", class(sets))
plot_scopus_intersections(sets, highlight = sets$label[3])
```

## Reading the fuller record

The Search API returns a few fields per record. To read the abstract and the
fuller metadata for a record you already know, `scopus_abstract()` calls the
Abstract Retrieval API, by DOI or 'Scopus' identifier. A batch is resilient, so
an identifier that cannot be found yields a row of `NA`s with a warning rather
than stopping the run.

```{r eval = FALSE}
ab <- scopus_abstract(c("10.1038/nature14539", "10.1103/PhysRevLett.116.061102"))
```

The result is a tibble of class `scopus_abstracts`, one row per identifier. To
show its shape without a key, here is a stand-in with the same columns. The
`abstract` column holds prose far too wide to typeset whole, so the columns are
listed by name and the identifying ones shown as a table.

```{r}
ab <- tibble::tibble(
  id          = c("10.1038/nature14539", "10.1103/PhysRevLett.116.061102"),
  scopus_id   = c("85060000001", "84960000002"),
  doi         = c("10.1038/nature14539", "10.1103/PhysRevLett.116.061102"),
  title       = c("Deep learning",
                  "Observation of gravitational waves from a binary black hole merger"),
  abstract    = c("Deep learning allows computational models that are ...",
                  "On 14 September 2015 the two detectors of LIGO observed ..."),
  publication = c("Nature", "Physical Review Letters"),
  year        = c(2015L, 2016L),
  citations   = c(42000L, 5400L)
)
class(ab) <- c("scopus_abstracts", class(ab))
names(ab)

ab[, c("title", "publication", "year", "citations")]
substr(ab$abstract[2], 1, 40)
```

## Beyond five thousand records

A single Search API query returns at most its first 5000 records under the
ordinary offset paging. When you need the whole of a larger result set in one
pass, `scopus_fetch(cursor = TRUE)` follows the API's cursor instead, which has no
such ceiling.

```{r eval = FALSE}
recs <- scopus_fetch("TITLE-ABS-KEY(microplastics)", cursor = TRUE)
nrow(recs)
#> [1] 38374
```

That query matched 38,374 records when this article was written (the literature
keeps growing), several times the offset ceiling, and the cursor retrieves all
of them in one call. The records then arrive in the API's deep-paging order
rather than sorted by relevance, which is the right trade for a complete
harvest. This is the one-call
alternative to the year-partitioned plan in the *Search plans and quota-aware
retrieval* article: a plan keeps each cell under the ceiling and preserves
relevance order, whereas `cursor = TRUE` harvests the whole set in a single pass.
Reach for the plan when you want cached, resumable cells, and the cursor when you
want the complete set at once.
