---
title: "Comparing topics over time"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing topics over time}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
# The plotting chunks below need ggplot2 (a suggested package); skip them
# gracefully when it is not installed.
has_ggplot2 <- requireNamespace("ggplot2", quietly = TRUE)
```

```{r setup}
library(scopusflow)
```

A common bibliometric question is not how large a literature is, but how its
internal emphasis shifts over time. Within research on language learning, say, is
the share of work that also discusses effect sizes growing? `scopus_compare_topics()`
answers exactly this, and `plot_scopus_comparison()` shows the answer. The
comparison itself contacts the API, so it is shown but not run; the plotting is
reproduced offline from an object of the same shape.

## What the comparison measures

For each year and each comparison term, the function counts the records matching
the reference topic *and* that term, and expresses it as a percentage of the
records matching the reference *alone*. A value of 30% for "effect size" in 2020
means that 30% of the language-learning records that year also mention effect
sizes. The reference is the denominator, so it sits at 100% by construction and
is not drawn.

```{r eval = FALSE}
cmp <- scopus_compare_topics(
  reference_query  = "language learning",
  comparison_terms = c("effect size", "Bayesian", "replication"),
  years            = 2014:2021,
  field            = "TITLE-ABS-KEY"
)
```

## The shape of the result

The result is a tidy table with one row per topic and year. We build one here
with the same columns so the rest of the article runs without a key.

```{r}
years <- 2014:2021
cmp <- tibble::tibble(
  query = "q",
  query_type = c(rep("reference", length(years)),
                 rep("comparison", 3 * length(years))),
  abridged_query = c(rep("language learning", length(years)),
                     rep(c("effect size", "Bayesian", "replication"),
                         each = length(years))),
  year = rep(years, 4),
  n = c(rep(900, length(years)),
        round(seq(120, 360, length.out = length(years))),
        round(seq(40, 200, length.out = length(years))),
        round(seq(20, 150, length.out = length(years)))),
  reference_n = rep(900, 4 * length(years)),
  comparison_percentage = c(rep(100, length(years)),
                            seq(13, 40, length.out = length(years)),
                            seq(4, 22, length.out = length(years)),
                            seq(2, 17, length.out = length(years))),
  average_comparison_percentage = c(rep(100, length(years)),
                                    rep(c(27, 13, 9), each = length(years)))
)
class(cmp) <- c("scopus_comparison", class(cmp))
cmp
```

The `comparison_percentage` column is the per-year share, and
`average_comparison_percentage` is the same ratio computed over the whole period,
which is what orders the topics. A year in which the reference has no records has
no defined share and is recorded as `NA` rather than as a misleading zero.

## A first plot

```{r eval = has_ggplot2, fig.alt = "Three topics' share of the language-learning literature from 2014 to 2021", fig.width = 7.5, fig.height = 4.5}
plot_scopus_comparison(cmp)
```

The chart uses whole-number year breaks, a colour-blind-safe palette and, because
there are only a few topics, labels the lines directly so the reader need not
match colours to a legend. Each label carries the topic's total record count.

## Drawing the eye to one topic

When one topic is the focus of a figure, `highlight` draws it in an accent colour
and greys the rest, which keeps the context visible without letting it compete.

```{r eval = has_ggplot2, fig.alt = "The same chart with the Bayesian topic highlighted against the others in grey", fig.width = 7.5, fig.height = 4.5}
plot_scopus_comparison(cmp, highlight = "Bayesian")
```

## Adjusting the labels

The count suffix on each label can be turned off when a cleaner look is wanted.

```{r eval = has_ggplot2, fig.alt = "The comparison chart without record counts in the labels", fig.width = 7.5, fig.height = 4.5}
plot_scopus_comparison(cmp, pub_count_in_legend = FALSE)
```

The return value is an ordinary [ggplot2](https://ggplot2.tidyverse.org) object,
so any further adjustment, a different theme or a saved file, is one `+` or one
`ggplot2::ggsave()` away.

## Reading the result as a table

Sometimes the numbers matter more than the picture. Because the output is a
tibble, the usual tools apply: here are the topics ranked by their average share.

```{r}
comp <- cmp[cmp$query_type == "comparison", ]
unique(comp[, c("abridged_query", "average_comparison_percentage")])
```
