---
title: "Comparing topics over time"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing topics over time}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
# Render data frames and tibbles as tables rather than console text.
local({
  kp <- function(x, ...) {
    if (any(vapply(x, is.list, logical(1)))) return(knitr::normal_print(x))
    knitr::knit_print(knitr::kable(x))
  }
  for (cls in c("data.frame", "tbl_df", "tbl")) {
    registerS3method("knit_print", cls, kp, envir = asNamespace("knitr"))
  }
})
# The plotting chunks below need ggplot2 (a suggested package); skip them
# gracefully when it is not installed.
has_ggplot2 <- requireNamespace("ggplot2", quietly = TRUE)
```

```{r setup}
library(scopusflow)
```

A common bibliometric question is not how large a literature is, but how its
internal emphasis shifts over time. Within deep-learning research, say, is the
share of work that also concerns medical imaging growing faster than the share
about computer vision? `scopus_compare_topics()` answers exactly this, and
`plot_scopus_comparison()` shows the answer. The comparison itself contacts the
API, so it is shown but not run; the plotting is reproduced offline from an object
of the same shape.

## What the comparison measures

For each year and each comparison term, the function counts the records matching
the reference topic *and* that term, and expresses it as a percentage of the
records matching the reference *alone*. A value of 30% for 'computer vision' in
2020 means that 30% of the deep-learning records that year also mention computer
vision. The reference is the denominator, so it sits at 100% by construction and
is not drawn.

```{r eval = FALSE}
cmp <- scopus_compare_topics(
  reference_query  = "deep learning",
  comparison_terms = c("computer vision", "natural language processing",
                       "medical imaging", "drug discovery"),
  years            = 2013:2021,
  field            = "TITLE-ABS-KEY"
)
```

## The shape of the result

The result is a tidy table with one row per topic and year. We build one here
with the same columns so the rest of the article runs without a key. The
reference set grows over the period, which the uncertainty band will reflect.

```{r}
years <- 2013:2021
ref_n <- round(seq(400, 1600, length.out = length(years)))
mk <- function(from, to) round(seq(from, to, length.out = length(years)))
counts <- list(
  "computer vision" = mk(140, 720),
  "natural language processing" = mk(90, 540),
  "medical imaging" = mk(15, 260),
  "drug discovery" = mk(8, 170)
)
cmp <- tibble::tibble(
  query = "q",
  query_type = c(rep("reference", length(years)),
                 rep("comparison", length(counts) * length(years))),
  abridged_query = c(rep("deep learning", length(years)),
                     rep(names(counts), each = length(years))),
  year = rep(years, length(counts) + 1),
  n = c(ref_n, unlist(counts, use.names = FALSE)),
  reference_n = rep(ref_n, length(counts) + 1),
  comparison_percentage = 100 * c(ref_n, unlist(counts, use.names = FALSE)) /
    rep(ref_n, length(counts) + 1),
  average_comparison_percentage = c(rep(100, length(years)),
                                    rep(c(40, 33, 15, 9), each = length(years)))
)
class(cmp) <- c("scopus_comparison", class(cmp))
cmp
```

The `comparison_percentage` column is the per-year share, and
`average_comparison_percentage` is the same ratio computed over the whole period,
which is what orders the topics. A year in which the reference has no records has
no defined share and is recorded as `NA` rather than as a misleading zero.

## A first plot

```{r eval = has_ggplot2, fig.alt = "Four application areas' share of the deep-learning literature from 2013 to 2021, with shaded uncertainty bands and an in-panel legend", fig.width = 8, fig.height = 4.6}
plot_scopus_comparison(cmp, legend_inside = TRUE)
```

Here `legend_inside = TRUE` places the topic key inside the panel, in whichever
corner has the most free space, rather than labelling each line at its end. Left
at its default the chart uses whole-number year breaks, a colour-blind-safe
palette and, because there are only a few topics, labels the lines directly so the
reader need not match colours to a legend. Each label carries the topic's total
record count. The
shaded band around each line is a Wilson stability range: it is wide in the early
years, when the reference set is small and the share would move easily, and
narrows as the literature grows. Because 'Scopus' returns exact counts rather than
a sample, the band is illustrative rather than a confidence interval, a point the
`plot_scopus_comparison()` help page sets out.

## When lines converge at the right end

Direct labels are legible only if they do not overlap, and topics sometimes end
the period at nearly the same share. `plot_scopus_comparison()` spreads
converging labels apart automatically, at the point the figure is actually
drawn, so they stay readable at any figure size rather than stacking into an
unreadable pile. Here six sub-areas of materials-science research all end
2013–2021 within three points of one another.

```{r}
years <- 2013:2021
ends <- c(18, 18.6, 19.2, 19.8, 20.4, 21)
names(ends) <- c("graphene", "perovskites", "MXenes", "COFs", "MOFs", "aerogels")
ref_n <- round(seq(500, 2000, length.out = length(years)))
converge <- function(end) round(end * (0.5 + 0.5 * (0:(length(years) - 1)) / (length(years) - 1)) * ref_n / 100)
counts <- lapply(ends, converge)

cmp_converging <- tibble::tibble(
  query = "q",
  query_type = c(rep("reference", length(years)),
                 rep("comparison", length(counts) * length(years))),
  abridged_query = c(rep("energy materials", length(years)),
                     rep(names(counts), each = length(years))),
  year = rep(years, length(counts) + 1),
  n = c(ref_n, unlist(counts, use.names = FALSE)),
  reference_n = rep(ref_n, length(counts) + 1),
  comparison_percentage = 100 * c(ref_n, unlist(counts, use.names = FALSE)) /
    rep(ref_n, length(counts) + 1),
  average_comparison_percentage = c(rep(100, length(years)),
                                    rep(ends, each = length(years)))
)
class(cmp_converging) <- c("scopus_comparison", class(cmp_converging))
```

```{r eval = has_ggplot2, fig.alt = "Six materials-science sub-areas converging to similar shares by 2021, with end labels automatically spread apart rather than overlapping", fig.width = 8, fig.height = 4.6}
plot_scopus_comparison(cmp_converging)
```

Without this, six labels ending within three points of each other would print
on top of one another; here every one is still readable, each still joined to
its own line by a short leader.

## Drawing the eye to one topic

When one topic is the focus of a figure, `highlight` draws it in an accent colour
and greys the rest, which keeps the context visible without letting it compete.

```{r eval = has_ggplot2, fig.alt = "The same chart with the medical-imaging topic highlighted against the others in grey", fig.width = 8, fig.height = 4.6}
plot_scopus_comparison(cmp, highlight = "medical imaging")
```

## Adjusting the labels

The count suffix on each label can be turned off, and the uncertainty band can be
removed, when a cleaner look is wanted.

```{r eval = has_ggplot2, fig.alt = "The comparison chart without record counts or bands", fig.width = 8, fig.height = 4.6}
plot_scopus_comparison(cmp, pub_count_in_legend = FALSE, interval = FALSE)
```

The return value is an ordinary [ggplot2](https://ggplot2.tidyverse.org) object,
so any further adjustment, a different theme or a saved file, is one `+` or one
`ggplot2::ggsave()` away.

## Reading the result as a table

Sometimes the numbers matter more than the picture. Because the output is a
tibble, the usual tools apply: here are the topics ranked by their average share.

```{r}
comp <- cmp[cmp$query_type == "comparison", ]
unique(comp[, c("abridged_query", "average_comparison_percentage")])
```
