---
title: "Search plans and quota-aware retrieval"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Search plans and quota-aware retrieval}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
# Render data frames and tibbles as tables rather than console text.
local({
  kp <- function(x, ...) {
    if (any(vapply(x, is.list, logical(1)))) return(knitr::normal_print(x))
    knitr::knit_print(knitr::kable(x))
  }
  for (cls in c("data.frame", "tbl_df", "tbl")) {
    registerS3method("knit_print", cls, kp, envir = asNamespace("knitr"))
  }
})
```

```{r setup}
library(scopusflow)
```

The Elsevier Scopus Search API is generous but bounded. A weekly quota limits how
many requests you may make, a short-term rate limit caps how fast you may make
them, and no single query will return more than its first 5000 records. This
article shows how scopusflow works within those bounds so that a large retrieval
is reproducible, efficient and resumable. The steps that contact the API need a
key and are not run here. Everything else runs offline.

## A query, built safely

Most queries combine a few terms under a field tag. `scopus_query()` assembles
them without the bracket and tag mistakes that creep in when fragments are pasted
together by hand.

```{r}
q <- scopus_query("language learning", "effect size", .field = "TITLE-ABS-KEY")
q
```

The recognised field tags, and what each one searches, are listed by
`scopus_field_tags()`.

```{r}
scopus_field_tags()
```

## Describing the search as a plan

A plan records exactly what will be fetched, so it can be saved, reviewed and
re-run. Partitioning by year is the recommended way to stay under the
5000-record ceiling, since each year becomes its own cell.

```{r}
plan <- scopus_plan(q, years = 2010:2020, partition = "year")
plan
```

Each cell carries the query, the year, the view and the page size. The page size
deserves a moment's attention, because it is where quota is won or lost.

## Why page size is a quota decision

Scopus charges quota per request, not per record. A page may hold up to 200
records under the `STANDARD` view, or 25 under `COMPLETE`. Retrieving a thousand
records in pages of 200 therefore costs five requests, where pages of 25 would
cost forty. For that reason `page_size` defaults to the largest the view allows,
which is the same efficiency `rscopus` relies on, and is in no sense an evasion
of the quota: every request is counted, and the 5000-record ceiling still holds.

```{r}
scopus_plan(q, view = "STANDARD")$page_size[1]
scopus_plan(q, view = "COMPLETE")$page_size[1]
```

## Sizing before spending

Counting is cheap and does not download records, so it is worth doing first. The
count comes back with the parsed quota attached, which lets a workflow decide
whether it has the allowance to proceed.

```{r eval = FALSE}
n <- scopus_count(q, years = 2010:2020)
n
attr(n, "quota")
```

That allowance is parsed from the response headers by `scopus_quota()`. To show
its shape without a network call, apply it to a constructed response:

```{r}
resp <- httr2::response(
  status_code = 200,
  headers = list(
    `X-RateLimit-Limit`     = "20000",
    `X-RateLimit-Remaining` = "19987",
    `X-RateLimit-Reset`     = "1700000000"
  )
)
scopus_quota(resp)
```

## Fetching, with caching and resume

`scopus_fetch_plan()` runs each cell in turn. Given a cache directory it writes
each cell to disk as it completes, so a run interrupted halfway, or stopped by
the quota, resumes from where it left off rather than paying for the same cells
again.

```{r eval = FALSE}
records <- scopus_fetch_plan(
  plan,
  cache_dir = scopus_cache_dir(),
  resume = TRUE
)
records
```

The cache lives under `scopus_cache_dir()`. To force a fresh retrieval, empty it
with `scopus_cache_clear()`. Both are shown but not run, so the article does not
touch a real cache:

```{r eval = FALSE}
scopus_cache_dir()    # where completed cells are written
scopus_cache_clear()  # remove them, so the next run re-fetches from scratch
```

The result is a `scopus_records` tibble, the same shape returned by
`scopus_fetch()` for a single query and by the bundled `example_records`.

```{r}
example_records
```

## Combining separate retrievals

Results gathered in separate runs combine safely with `scopus_combine()`, which
renumbers the records and can drop duplicates by Scopus identifier or DOI. This
is preferable to `rbind()`, which would leave duplicate entry numbers.

```{r}
scopus_combine(example_records, example_records, dedupe = TRUE)
```

## When the ceiling bites

A query matching more than 5000 records cannot be retrieved in full from a single
call. `scopus_fetch()` returns the first 5000 and warns. The remedy is to split
the search by year, or by any other facet, so that each cell stays under the
ceiling. `scopus_count()` tells you in advance whether a split is needed.

## Handling interruptions

Network and API problems are raised as typed conditions, all inheriting from
`scopus_error`, so a long retrieval can respond to them rather than stopping
dead.

```{r eval = FALSE}
result <- tryCatch(
  scopus_fetch_plan(plan, cache_dir = scopus_cache_dir()),
  scopus_error_rate_limit = function(e) {
    message("Rate limited; the cached cells are safe. Try again later.")
    NULL
  }
)
```

Because each completed cell is already cached, resuming after such a pause costs
nothing for the work already done.
