---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
# Render data frames and tibbles as tables rather than console text.
local({
  kp <- function(x, ...) {
    if (any(vapply(x, is.list, logical(1)))) return(knitr::normal_print(x))
    knitr::knit_print(knitr::kable(x))
  }
  for (cls in c("data.frame", "tbl_df", "tbl")) {
    registerS3method("knit_print", cls, kp, envir = asNamespace("knitr"))
  }
})
```

```{r setup}
library(scopusflow)
```

This vignette is fully reproducible without a Scopus API key. It draws on a small
static fixture bundled with the package, so the whole workflow can be shown
offline. The few steps that genuinely need the API are shown but not run.

## Describing a search as a plan

A plan separates describing a search from executing it. Plans are inspectable,
saveable and version-controllable, and they can be partitioned, for example by
year, so that a large retrieval stays under the API's `start < 5000` ceiling and
can be cached and resumed.

```{r}
plan <- scopus_plan(
  "machine translation",
  years     = 2018:2020,
  field     = "TITLE-ABS-KEY",
  partition = "year"
)
plan
```

Each row is one query cell. Field tags wrap the query and years become a date
filter:

```{r}
scopus_plan("language learning", field = "TITLE")$query
scopus_plan("x", years = 2015:2020)$date

# A plan is a classed object; is_scopus_plan() confirms it.
is_scopus_plan(plan)
```

## Sizing and fetching

`scopus_has_key()` reports whether a key is configured, without revealing it. It
is the guard the package's own examples use to skip the steps that need the API,
so it is the natural switch for a reproducible script:

```{r}
scopus_has_key()
```

With a key configured, you size a search cheaply and then execute the plan,
optionally caching each cell so that an interrupted run resumes without
re-spending quota. These contact the API, so they are not evaluated here:

```{r eval = FALSE}
scopus_count("machine translation", years = 2018:2020, field = "TITLE-ABS-KEY")

records <- scopus_fetch_plan(plan, cache_dir = scopus_cache_dir(), resume = TRUE)
```

## The record schema

Whether records come from the API or from the bundled example data, they share
one stable schema. The package ships a small, already normalised set, which we
use here to continue offline:

```{r}
records <- example_records
records

# A record set is a classed tibble; is_scopus_records() confirms the contract.
is_scopus_records(records)
```

`scopus_records()` produces this same shape from a raw API response, flattening
the nested result into one row per record:

```{r}
# A couple of records in the shape the Scopus API returns them.
raw <- list(entry = list(
  list(`dc:identifier` = "SCOPUS_ID:1", `prism:doi` = "10.1000/aaa",
       `dc:title` = "A reproducible workflow", `dc:creator` = "Smith J.",
       `prism:publicationName` = "J. Bibliometrics",
       `prism:coverDate` = "2020-05-01", `citedby-count` = "12"),
  list(`dc:identifier` = "SCOPUS_ID:2", `prism:doi` = "10.1000/bbb",
       `dc:title` = "Quota-aware querying", `dc:creator` = "Doe A.",
       `prism:publicationName` = "Scientometrics Today",
       `prism:coverDate` = "2021-01-10", `citedby-count` = "3")
))
scopus_records(raw, query = "TITLE-ABS-KEY(workflow)")
```

## DOIs and change tracking

Extract a clean, deduplicated DOI list for import into a reference manager, and
compare two retrievals to see exactly what changed:

```{r}
dois <- scopus_extract_dois(records)
dois

# Suppose a later retrieval added one DOI and dropped another.
later <- c(dois[-1], "10.1000/example.999")
scopus_diff_dois(old = dois, new = later)
```

You can write the DOIs to a path you specify, and read the file back to see
exactly what lands on disk:

```{r, comment = "", collapse = FALSE}
out <- file.path(tempdir(), "dois.csv")
scopus_extract_dois(records, file = out)
writeLines(readLines(out))
```

## Comparing topic trends

`scopus_compare_topics()` measures how the internal emphasis of a literature
shifts, expressed as each comparison topic's yearly share of the reference
literature. It issues one count request per term per year, so it needs the API:

```{r eval = FALSE}
cmp <- scopus_compare_topics(
  reference_query  = "language learning",
  comparison_terms = c("effect size", "Bayesian"),
  years            = 2015:2020,
  field            = "TITLE-ABS-KEY"
)
plot_scopus_comparison(cmp)
```

The result is a tidy table with one row per topic and year, which
`plot_scopus_comparison()` draws with direct line labels, a colour-blind-safe
palette and shaded stability bands. `vignette("comparing-topics")` builds the
object offline, shows the plot in its variations and explains how to read the
bands.

## Author keywords and references

A search only returns the fields the Search API carries. Author keywords and a
document's own reference list need `view = "COMPLETE"` and Abstract Retrieval
respectively, both at a materially different quota cost from an ordinary
search; `vignette("keywords-and-references")` walks through both, and
`scopus_corpus()`, which combines them into a minimal
`id`/`title`/`year`/`keywords`/`references` shape for downstream tools.

## Export and interoperability

Hand results to `bibliometrix`-style workflows, or save and reload them:

```{r}
head(as_bibliometrix(records))

path <- file.path(tempdir(), "records.rds")
write_scopus_records(records, path)
identical(read_scopus_records(path), records)
```

## Handling failures

Network and API problems surface as typed conditions, all inheriting from
`scopus_error`, so a workflow can respond to them in code:

```{r eval = FALSE}
tryCatch(
  scopus_fetch("..."),
  scopus_error_no_key     = function(e) message("No API key configured."),
  scopus_error_rate_limit = function(e) message("Rate limited, so backing off."),
  scopus_error            = function(e) message("Scopus error: ", conditionMessage(e))
)
```
