---
title: "Tracking how a literature changes between retrievals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tracking how a literature changes between retrievals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
# Render data frames and tibbles as tables rather than console text.
local({
  kp <- function(x, ...) {
    if (any(vapply(x, is.list, logical(1)))) return(knitr::normal_print(x))
    knitr::knit_print(knitr::kable(x))
  }
  for (cls in c("data.frame", "tbl_df", "tbl")) {
    registerS3method("knit_print", cls, kp, envir = asNamespace("knitr"))
  }
})
```

```{r setup}
library(scopusflow)
```

A literature is a moving target. Run the same search a few months apart and the
result will have grown, and perhaps lost a record that was re-indexed. This
article shows how to see exactly what changed and how to merge retrievals safely.
It runs offline on the bundled `example_records`, a corpus of 138 real journal
articles the package ships because 'Scopus' records may not be redistributed.
That corpus is a complete harvest of one query from 2015 to 2024, so a pull that
stopped at 2023 and a later one that reaches 2024 are both genuine slices of the
same search.

## The baseline

The first retrieval ran at the end of 2023 and returned everything published up
to then.

```{r}
baseline <- example_records[example_records$year <= 2023, ]
nrow(baseline)
```

## A later retrieval

A year on, the search is repeated. It now picks up the 2024 papers, and one
record that was present the first time has since been re-indexed and no longer
matches.

```{r}
later <- example_records[-1, ]
nrow(later)
```

## What changed

`scopus_diff_dois()` reports which DOIs were added, removed or unchanged between
the two retrievals, and prints the counts in each category.

```{r}
changes <- scopus_diff_dois(old = baseline, new = later)
print(changes)
```

The newly indexed papers come back as `added`, the records present both times as
`unchanged`, and anything dropped from the later pull as `removed`. The counts
work out at fourteen added, one removed and 112 unchanged: fourteen because that
is how many of the 2024 papers carry a DOI, and 112 rather than 113 because the
re-indexed record is no longer among them. Records without a DOI cannot be
tracked this way at all, which is one reason to prefer the 'Scopus' identifier
when there is one.

To act on one category, filter the table, which is an ordinary tibble.

```{r}
head(changes[changes$status == "added", ])
```

## Merging without duplicates

To keep a cumulative set across retrievals, combine them. `scopus_combine()`
renumbers the records and, with `dedupe = TRUE`, keeps each one once by 'Scopus'
identifier or DOI, so the records the two pulls share are not doubled.

```{r}
combined <- scopus_combine(baseline, later, dedupe = TRUE)
nrow(combined)
```

That is 149 rows for 138 distinct articles, and the gap is instructive. These
records carry no 'Scopus' identifier, not having come from 'Scopus', so
de-duplication falls back to the DOI. The eleven that arrived without one have
no key to match on, and so survive in both copies. A live harvest carries an
identifier on every record, so the same call on two real pulls returns each
article once.

The base `c()` method concatenates record sets directly, renumbering but without
de-duplicating, so it is the building block that `scopus_combine()` adds the
duplicate handling to.

```{r}
stacked <- c(baseline, later)
nrow(stacked)
```

## Keeping a record of each pull

Saving each retrieval lets you compare against it next time. The `.rds` form
round-trips exactly.

```{r}
path <- file.path(tempdir(), "baseline.rds")
write_scopus_records(baseline, path)
identical(read_scopus_records(path), baseline)
```

In a live setting the later retrieval would come from the API rather than from a
slice of the bundled corpus, with everything else unchanged.

```{r eval = FALSE}
later <- scopus_fetch("graphene supercapacitor", field = "TITLE-ABS-KEY")
scopus_diff_dois(old = read_scopus_records(path), new = later)
```
