---
title: "Using the code-free app"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the code-free app}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = FALSE, comment = "", eval = FALSE)
# Console colour carries no meaning on a rendered page. pkgdown turns it on for
# its own build, and the escape sequences then reach the reader as literal text,
# so colour is switched off here for a plain vignette render and a site build
# alike. The fixed width keeps tibbles inside the documentation column.
options(cli.num_colors = 1, cli.hyperlink = FALSE, crayon.enabled = FALSE,
        width = 80)
```

<!-- The screenshots in figures/ were captured from run_app() with Demo mode
     left on, so they need no key and no network. Recapture them whenever the
     app's layout changes. Unlike the figures in the other articles, which are
     drawn when the vignette is knitted, a static image of an interface drifts
     silently. -->


Not every search starts in an editor. `run_app()` opens a local Shiny app that
drives the whole scopusflow workflow through a browser tab, from describing a
search to exporting the records, without writing any R. It runs on your own
machine, so your Scopus key never leaves it and requests come from your own
network. The app doubles as a way into the package itself, because it mirrors
every choice you make back as a runnable R script. Anything you can do in the
app you can also do from code, and the script panel shows you how.

This article describes the app without running it, since that would need a live
server, so the code chunks are shown but not evaluated. The screenshots were
taken from the running app with Demo mode left on, which needs no key and no
network, so everything they show can be reproduced on your own machine in a
minute.

## Launching the app

The app needs the suggested packages shiny, bslib and callr, and ggplot2 for the
plots. With those installed, one call opens it.

```{r}
library(scopusflow)
run_app()
```

By default it listens only on `127.0.0.1`, so it is reachable from your own
machine alone, and it opens your browser at the address it prints. Passing
`launch.browser = FALSE` leaves the browser to you, and `port` fixes the port
when you need a stable address.

## Demo mode

Demo mode is switched on when the app opens, so you can walk the whole flow with
no key at all. With it on, the harvest is simulated, the live terminal still
streams its per-cell progress, the tables and plots still render, and the Compare
topics tab still draws a figure. It is the quickest way to learn where each
control sits before you spend any quota. The records come from `example_records`,
the corpus of 138 real journal articles the package bundles because Scopus
records may not be redistributed, so every panel is exercised on real titles,
DOIs, journals and citation counts, and the by-year chart shows a real
publication curve with the shape the literature actually has. That corpus spans
2015 to 2024, and the year slider opens on exactly that span, so the first run
has records for every cell it announces. A year outside the span is drawn from
the nearest year the corpus holds, and never twice, so a range of 2020 to 2026
replays what 2020 to 2024 does; the live terminal names each cell that falls
outside as it passes. Only the topic comparison is synthesised, since it counts
whole literatures, which no set of records in hand can answer for. When you are
ready for real results, paste your Scopus key into the field at the top of the
sidebar and switch Demo mode off. The key stays in the running session and is
never written into the generated script.

## Describing and sizing a search

The sidebar is where you describe the search. You enter your terms, choose which
field to search in, set a year range, and pick the level of detail. Partitioning
by year is recommended and switched on by default, because it keeps each cell
under the API's offset ceiling, the same reasoning behind a partitioned
`scopus_plan()` in code. Check size runs a cheap count first, the same one
`scopus_count()` performs, so you can see how much a query would retrieve before
committing to it. In demo mode it reports instead how many records it would draw
from the bundled corpus. Max records per cell caps what each cell brings back and
is pre-filled with 200, so a search that counts in the thousands still returns at
most 200 records a cell until you raise it or clear the field; the size note says
what the cap would leave you with, since a cell stopped at a cap you set is
short by request and passes without a warning.

![The scopusflow sidebar, with a Scopus API key field, a Demo mode checkbox switched on, the search terms "graphene supercapacitor", a Search in selector reading "Title, abstract, keywords", a Partition by year checkbox, a year-range slider set to 2021 to 2026, a STANDARD or COMPLETE detail choice, a maximum records per year field, and the Check size, Fetch records and Cancel buttons.](figures/app-sidebar.png)

## Fetching, with a live terminal

Fetch records starts the harvest in a background process, so the app stays
responsive while it works. The Live terminal panel tails the worker's output,
streaming a line as each year-cell completes, and a progress bar tracks how far
through the plan the run has reached. Cancel ends the background process at
once, which does abandon whatever cell was in flight. Every cell that had
already finished was checkpointed as it completed, and a checkpoint is written
to a temporary file and renamed into place, so a cancelled run leaves the cache
whole and a later run resumes from it. Internally the app builds a
`scopus_plan()` from your choices and hands it to `scopus_fetch_plan()` with a
cache directory and resume turned on, so an interrupted or quota-limited run
picks up where it left off.

![The Retrieval panel part-way through a demo-mode harvest. A note reads "Demo plan: 6 year-cells; would draw 57 records from the bundled corpus", a label reads "Fetching cell 4 of 6" above a progress bar at 50%, and the expanded Live terminal below shows four lines, one for each cell the run has reached, each reading "Cell n/6: demo records for graphene supercapacitor" with the year.](figures/app-terminal.png)

## The reproducible code panel

The Reproducible code panel is what turns the app into a tutorial. It mirrors
every choice you make, the query, the field, the years, whether you partition by
year, and the Compare topics options, and rewrites a runnable R script as you go.
Change a control and the script updates at once, so you can see exactly which
argument each choice sets. The key never appears in the script. The panel notes
that it is read from the `SCOPUS_API_KEY` environment variable, so a script you
share carries the method but not your credentials. Every script opens with the
package version that wrote it, and one written in demo mode says so at the top,
so a file downloaded from a session that retrieved nothing does not read as the
provenance of a real harvest. Download script saves exactly what you see.

![The Reproducible code panel, holding the script the app has written from the choices in the sidebar: a library call, a scopus_plan() with the query, years, field and yearly partition, a scopus_count() to size the search, a scopus_fetch_plan() with a cache directory and resume set, and scopus_top() and autoplot() calls to inspect the result. A comment records that the key is read from the SCOPUS_API_KEY environment variable and never written into the script.](figures/app-code.png)

## Reading the results

When the run finishes, the Results tabs come to life. Records shows a table of
titles, years, sources and citations. It says how many records the run returned
and, above a harvest of more than twenty-five, that the table is a preview of
them, since the exports below carry the whole set. By year draws the
publications-per-year trend through `autoplot()`. Top sources and Top authors
tally the most frequent of each with `scopus_top()` and `plot_scopus_top()`.
Every figure is drawn by the same library functions you would call from a script,
so the app shows you nothing you could not reproduce.

![The Results panel with tabs for Records, By year, Top sources, Top authors, Compare topics and Export. The Records tab is open, showing the first rows of a table of titles, years, journals and citation counts from the demo harvest. The first row reads "Recent trends in graphene supercapacitors: from large area to microsupercapacitors", published in Sustainable Energy & Fuels in 2021 with 211 citations, and the rows below it are likewise real articles from the bundled corpus.](figures/app-results.png)

## Comparing topics

The Compare topics tab asks a different question from a harvest. It retrieves no
records at all, and measures instead how a set of sub-topics co-occur with your
search over time, as a share of it, with your search terms acting as the
reference topic. You enter comma-separated comparison terms, optionally pick one
to highlight, and toggle the stability band and whether record counts appear on
the labels. Because each term needs one count request per year, and the reference
topic one more, the tab shows how many requests a comparison will make and warns
when the grid grows large. A comparison always covers a span of years: the one
on the slider, or the last six when Partition by year is off, and the request
note names it, since the sidebar then shows no years at all. In demo mode the
comparison is synthesised so you can see the figure offline, and a note beneath
it says so, since the figure's own caption names the Search API as every real
comparison's source. With a key it calls `scopus_compare_topics()` and draws the
result with `plot_scopus_comparison()`, the same figure the package produces,
and a Comparison download saves the underlying numbers as a CSV.

## Exporting

Every result the app shows comes with one-click export. The records can be saved
as an RDS file, as a clean DOI list, and as BibTeX and RIS for a reference
manager such as Zotero or EndNote, drawn from `as_bibtex()` and `as_ris()`, and
as a PRISMA-S search record in Markdown, drawn from `scopus_search_report()`. In
demo mode that record carries no plan or retrieval time, and says so, since the
records on screen were replayed from the bundled corpus and never retrieved. The
comparison can be saved as a CSV. None of this contacts the API again, because
it works on results already in hand. Between the downloaded script and the
exported records, a session in the app leaves you with both the data and the
code that produced it.

## Beyond the app

Anything you can do in the app you can also do from code, but the reverse does
not hold. The app covers planning, retrieval and the trend, top and comparison
analyses, and the package goes further. `scopus_intersections()` counts how a
set of named concepts and their intersections overlap in the literature, and
`plot_scopus_intersections()` draws the result. `scopus_abstract()` retrieves
abstract text and richer metadata than the Search API returns, and
`scopus_corpus()` builds on it to give a harvest author keywords and structured
references for tools further downstream. When a session in the app has taken
you as far as it goes, the downloaded script is the place to carry on from.
